---
title: "July 15 Quiz: POLS 3316"
output: html_notebook
---

## Quiz 2 and Problem Set 1 combined

### Due 11:59 PM July 16

## Part 1 - Complete before leaving class Friday

**Run code in R Studio or R Studio Cloud. Download the finished HTML or PDF file and upload to Blackboard by the due date for full credit**

**The areas with ``` followed by {r} and ending with ``` are code chunks. The Run dropdown arrow in the upper right of this box give you options for running the code. Run *Current Chunk* runs just the piece you are working on. After you have finished, you can *Run all* then save the result. Where you see Preview (above just left of center) there is a dropdown arrow. If you select "Knit to HTML" or "Knit to PDF", you can produce a copy that can be downloaded and submitted in Blackboard.**

**IF all else fails, screenshot your work and submit that in Blackbaord.**

For more help with this RMarkdown notebook here are some resources:

Cheatsheet - https://rmarkdown.rstudio.com/lesson-15.HTML

Code Chunks - https://rmarkdown.rstudio.com/lesson-3.html


1. a. (bonus) Create a vector named *my_vector* containing the sequence (1,2,3,4,5). Find the sum using the vector. 


```{r}



```

1. Write the code to create an object named *object_sum* whose value is the sum of the sequence of numbers (1,2,3,4,5) from problem 1 and then return the value of that object to the console. (*Hint: you'll use the assignment operator to assign an operation on the right side to object_sum on the left side. Hint2: You can type in a math operation or you can use your code from question 1. Hint 3: To return the value, I've already written the code, all you have to do is uncomment it.*)

```{r}

#object_sum 

```

2. What is the sum of the sequence of numbers (1,2,3,4,5)? (You can code in R for practice by pressing CTRL-ALT-I to create a code "chunk" in the answer spot, or you can add the answer and type it.)

Answer: 

3. How many observations (n) are in the sequence (1,2,3,4,5)?

Answer: 

3. a. Write simple code to create an object called *object_n* and assign this number to it. (*Hint: if you use more than the object name, the assignment operator, and a single digit number you are working too hard!*)

```{r}



```

4. Write the code to create a new object named *object_mean* whose value is the mean of the above values using *object_sum* and the number of observations *object_n*? (Hint: You'll use the assignment operator with the *object_mean* on the left side and a math operation using the other objects on the right side.)

```{r}



```


5. Get help with the *mean()* function in R in the next code chunk. If you need help for problem 6, you can type the same thing in the Console (bottom left): 

```{r}



```

6. Now write the code to find the mean of the above values using the *mean()* function in R? (*Hint: If you did the bonus question, 1a it will save you time here*)


```{r}



```

7. Write the code to find the sample variance of the above observations using the objects you created (*object_n, object_mean, object_sum*) and the sequence of values? (*Hint: Creating an object with the sample variance will help with question 8 but remember to return the object's value if you do.*)


```{r}



```

8. Write code to find the standard deviation? 

```{r}



```

9. What is the median of sequence of values from question 1. 

```{r}



```

10. What is the mode of  0, 1, 1, 2, 3, 5, 8, 13, 21, 34. 

```{r}



```

11. What is your favorite food?  

Answer: 

12. What is your Github username?  

Answer: 

13. What is the name of your project data set? 

Answer: 

14. What is the source of your data set?  

Answer: 

15. What would you like to learn from the data set in one or two sentences? (How to use data/R studio/statistitcs is an acceptable answer, but if you want to know more I can help you.)

Answer:

16. (Optional) How many variables does your data have? How many observations does your data have?

Answer:

17. (Optional) Does your data have observations across multiple time periods?

Answer:

### Part 2

*For the remaining problems, you can use math operations or R functions. You can use the methods above (creating objects, etc.) or you can just run the functions on the raw numbers. *

18. What is the sum of the sequence (5,10,15,20,25,30,35,40)

```{r}



```

19. How many observations (n) are in the sequence in question 18?

20. What is the mean of the sequence in question 18?

```{r}



```

21. What is the sample variance of the sequence in question 18?

```{r}



```

22. What is the sample standard deviation of the sequence in question 18?

```{r}



```

23. What is the median of the sequence in question 18?

24. What is the difference between a sample variance and a population variance? (One or two sentences.)

25. What is the sum of the sequence (3,7,5,5,6,7,7,9)

```{r}



```

26. What is the mode of the sequence?

27. How many observations (n) are in the sequence?

28. What is the mean of the sequence?

```{r}



```
