---
title: "July 15 Quiz: POLS 3316"
output: html_notebook
---

## Quiz 2 and Problem Set 1 combined

### Due 11:59 PM July 16

## Part 1 - Complete before leaving class Friday

**Run code in R Studio or R Studio Cloud. Download the finished HTML or PDF file and upload to Blackboard by the due date for full credit**

**The areas with ``` followed by {r} and ending with ``` are code chunks. The Run dropdown arrow in the upper right of this box give you options for running the code. Run *Current Chunk* runs just the piece you are working on. After you have finished, you can *Run all* then save the result. Where you see Preview (above just left of center) there is a dropdown arrow. If you select "Knit to HTML" or "Knit to PDF", you can produce a copy that can be downloaded and submitted in Blackboard.**

**IF all else fails, screenshot your work and submit that in Blackbaord.**

For more help with this RMarkdown notebook here are some resources:

Cheatsheet - https://rmarkdown.rstudio.com/lesson-15.HTML

Code Chunks - https://rmarkdown.rstudio.com/lesson-3.html


1. a. (bonus) Create a vector named *my_vector* containing the sequence (1,2,3,4,5). Find the sum using the vector. 


```{r}
my_vector <- 1:5


```

1. Write the code to create an object named *object_sum* whose value is the sum of the sequence of numbers (1,2,3,4,5) from problem 1 and then return the value of that object to the console. (*Hint: you'll use the assignment operator to assign an operation on the right side to object_sum on the left side. Hint2: You can type in a math operation or you can use your code from question 1. Hint 3: To return the value, I've already written the code, all you have to do is uncomment it.*)

```{r}
my_vector <- 1:5
print(my_vector)
object_sum <- sum(my_vector)
print(object_sum)

```

2. What is the sum of the sequence of numbers (1,2,3,4,5)? (You can code in R for practice by pressing CTRL-ALT-I to create a code "chunk" in the answer spot, or you can add the answer and type it.)

Answer: 
```{r}

print(object_sum)
```

3. How many observations (n) are in the sequence (1,2,3,4,5)?

Answer: 
```{r}
print("an observation is a data point, so there are 5 observations or data points")
print(my_vector)
```

3. a. Write simple code to create an object called *object_n* and assign this number to it. (*Hint: if you use more than the object name, the assignment operator, and a single digit number you are working too hard!*)

```{r}
object_n <- 5


```

4. Write the code to create a new object named *object_mean* whose value is the mean of the above values using *object_sum* and the number of observations *object_n*? (Hint: You'll use the assignment operator with the *object_mean* on the left side and a math operation using the other objects on the right side.)

```{r}
object_mean <- mean(object_n, object_sum)
print(object_mean)


```


5. Get help with the *mean()* function in R in the next code chunk. If you need help for problem 6, you can type the same thing in the Console (bottom left): 

```{r}

print(mean(my_vector))


```

6. Now write the code to find the mean of the above values using the *mean()* function in R? (*Hint: If you did the bonus question, 1a it will save you time here*)


```{r}
print(mean(my_vector))


```

7. Write the code to find the sample variance of the above observations using the objects you created (*object_n, object_mean, object_sum*) and the sequence of values? (*Hint: Creating an object with the sample variance will help with question 8 but remember to return the object's value if you do.*)


```{r}
object_variance <- c(object_n, object_mean, object_sum, my_vector)
print(object_variance)
print(var(object_variance))


```

8. Write code to find the standard deviation? 

```{r}

print(object_variance)
print(sd(object_variance))

```

9. What is the median of sequence of values from question 1. 

```{r}
object_median <- median(my_vector)
print(object_median)

```

10. What is the mode of  0, 1, 1, 2, 3, 5, 8, 13, 21, 34. 

```{r}
kat_mode <- function(x){
        #find unique numbers
        uniqx <- unique(x)
        #tabulate returns a vector of occurrences with index as a key
        uniqx[which.max(tabulate(match(x, uniqx)))]
        #use which.max to return largest number of occurrences
}

kat_vec <- c(0, 1, 1, 2, 3, 5, 8, 13, 21, 34)
result <- kat_mode(kat_vec)
print(result)


```

11. What is your favorite food?  

Answer: My favorite food is probably ice cream.

12. What is your Github username?  

Answer: shutup_kat

13. What is the name of your project data set? 

Answer: pols3316-summer2022

14. What is the source of your data set?  

Answer: github/pols3316-summer2022/problem sets

15. What would you like to learn from the data set in one or two sentences? (How to use data/R studio/statistics is an acceptable answer, but if you want to know more I can help you.)

Answer: Learn how to us R of course, but also how to find trends in data that can be helpful in predicting future trends. 

16. (Optional) How many variables does your data have? How many observations does your data have?

Answer: 10

17. (Optional) Does your data have observations across multiple time periods?

Answer: Im actually not sure what that means.

### Part 2

*For the remaining problems, you can use math operations or R functions. You can use the methods above (creating objects, etc.) or you can just run the functions on the raw numbers. *

18. What is the sum of the sequence (5,10,15,20,25,30,35,40)

```{r}

my_vec <- c(5,10,15,20,25,30,35,40)
print(sum(my_vec))

```

19. How many observations (n) are in the sequence in question 18?
```{r}
print(unique(my_vec))
print(" the number of observations in this set is: ")
print(length(unique(my_vec)))
```

20. What is the mean of the sequence in question 18?

```{r}

print(mean(my_vec))

```

21. What is the sample variance of the sequence in question 18?

```{r}

print(var(my_vec))

```

22. What is the sample standard deviation of the sequence in question 18?

```{r}

print(sd(my_vec))

```

23. What is the median of the sequence in question 18?
```{r}
print(median(my_vec))
```

24. What is the difference between a sample variance and a population variance? (One or two sentences.)

Answer: Population variance is a set of data coming from a population, while sample variance is a set of data from a specific sample of objects. A sample can be from a portion of the population or from something entierly different. 

25. What is the sum of the sequence (3,7,5,5,6,7,7,9)

```{r}
newvec <- c(3,7,5,5,6,7,7,9)
print(sum(newvec))


```

26. What is the mode of the sequence?
```{r}
#from function declared in problem 10
newresult <- kat_mode(newvec)
print(newresult)
```

27. How many observations (n) are in the sequence?
```{r}
print(unique(newvec))
print(" the number of observations in this set is: ")
print(length(unique(newvec)))
```

28. What is the mean of the sequence?

```{r}
print(mean(newvec))


```
